Environmental warming increases the absolute and relative importance of high-turnover energy channels in stream food webs

### 2022-11-18 doc/ms/Junker_temp-energy-flux_ms.Rmd
push before coauthor comments

@@

Warming temperatures are altering communities and trophic networks globally. While the influence of warming on food webs is often context-dependent, increasing temperatures are predicted to influence communities in two fundamental ways: 1) by reducing average body size and 2) by increasing individual metabolic rates. These warming-induced changes have the potential to influence the distribution of food web fluxes, food web stability, and the importance of environmental and biological factors in community assembly. Here, we quantified patterns and the relative distribution of organic matter fluxes through stream food webs spanning a natural ~25$^\circ$C temperature gradient. We then related these patterns to species- and community-level trait distributions (body size, *M* and biomass turnover, *P:B*) within and across ecosystems. We predicted that 1) communities in warmer streams would skew towards smaller *M* and higher *P:B* and 2) organic matter fluxes within warmer communities would increasingly skew towards smaller, higher *P:B* populations. Across the temperature gradient, communities in warmer streams were indeed composed of smaller-bodied and higher biomass-turnover populations on average. Additionally, material fluxes within warmer streams were increasingly skewed towards high-turnover populations, demonstrating that warming can restructure organic matter fluxes in both an absolute and relative sense. With warming, the relative distribution of organic matter fluxes also appeared to be increasingly 'non-random', suggesting the potential for stronger selection for these traits with increasing temperature. The emerging picture is a warmer world that is both smaller and faster. Our study suggests that a warming world will be 'smaller and faster', and that warming has the potential to shape pathways of energy and material flows in food webs in ways that may be predictable.

@@

<!--- warming and effects in food webs --->
<!--- Speeding up of food web: Species levels--->
Increasing global temperatures can influence the provision and maintenance of ecosystem services by modifying the network of species interactions that underpin ecosystem functions [@brose2012; @woodward2010; @deruiter1995; @thompson2012]. The effects of warming permeate across levels of biological organization from its control on individual metabolic rates [@gillooly2001; @brown2004] and biological activity (e.g., attack rate, handling time, growth rates, etc.; @dell2014, to broad-scale shifts in community assembly and structure [@saito2021; @gibert2019a; @nelson2017a]. Across global climate gradients, these temperature-induced changes have the potential to mediate food web stability [@baiser2019] by altering the acquisition and allocation of resources [@zhang2017] and the magnitudes and relative distribution of energy fluxes in food webs[@may1972; @mccann1998; @barnes2018].

Understanding how warming alters energy fluxes in food webs requires information about how temperature modifies connections between ecosystem structure (i.e. biodiversity) and function (BEF). Such connections hinge upon how warming influences functional trait distributions and how these functional traits translate to energy demands [@loreau2001; @norberg2001; @norberg2004]. Additionally, warming may modify both deterministic [e.g., environmental/niche filtering, @whittaker1962] and stochastic [e.g., neutral theory, @hubbell2001] community assembly processes [@saito2021], leading to many possible relationships between warming and species' abundance and trait distributions. For example, although temperature appears to have minimal control on species richness across taxonomic groups [e.g., @bastazini2021], warming may have profound effects on the dominance structure (i.e. evenness) of ecological communities, often by reducing community evenness and favoring a few well-adapted species [@hillebrand2008]. This strong environmental filtering is likely to skew trait distributions, to some extent, in natural communities [@therriault1999]. However, the relative distribution of traits can be modified by additional processes beyond environmental filtering that alter species relative abundance distributions [e.g., species interactions, @therriault1999; demographic stochasticity, @hubbell2001]. These additional processes can exaggerate or counter any skew in trait distribution imparted through environmental filtering and may modify the relationship between species' traits and the absolute and relative energy demands in food webs.

<!--- Connect traits to the 'speed' of ecosystem energy fluxes--->
Body size is a fundamental trait that is strongly influenced by temperature and has great potential to influence energy flux [@atkinson1994; @gardner2011; @daufresne2009]. Body size reductions arise from multiple mechanisms acting at different levels of organization, including the increased relative abundance of smaller species within warmer communities [@bergmann1848], smaller individuals in warmer populations [@james1970], and/or reduced body size of warmer individuals [@atkinson1994]. These temperature-size relationships are likely to influence the distribution of body sizes both across and within ecosystems, such that warmer communities should contain smaller species and populations, on average, and smaller organisms should be more dominant within a community. These changes can have important implications for ecosystems pattern and processes that control energy and material fluxes in ecosystems because body size is a strong determinant of an species life history patterns [@peters1983; @zeuss2017; @altermatt2010; @nelson2020b] and,developmental [@angilletta2004] or metabolic rates [@brown2004; @gillooly2001].

<!--- Write the main goals of the paragraph--->
Metabolic rate, in addition to being controlled by body size, is another key trait that is influence by temperature and tied to variation in energy flux through food webs [@gillooly2001]. Warming can influence metabolic rates indirectly through reductions in body size, as well as directly through its effects on subcellular kinetics [@bideault2019; @osmond2017]. These processes have been shown to modify ecosystem patterns through changes in population abundance [@bernhardt2018], consumer-resource interactions [@bideault2019], and food web structure [@gibert2019a]. Importantly, the kinetic effects of temperature are tied to many biological processes [@dell2014], among them growth rate [@gillooly2001], developmental rate [@zuo2012; @nelson2020b], voltinism [@zeuss2017], and population turnover rate [@huryn2007; @brown2004]. Taken together, these changes suggest that warming shoudl lead to a smaller, faster world in which ecosystem processes 'speed up' through the effects of smaller body size and higher turnover rates, with a potentially strong imprint on food web structure and energy fluxes [@gibert2019a].

We quantified patterns of organic matter flux in stream food webs across a natural temperature gradient (\~5--28$^\circ$C) in southwestern Iceland. Previous research in these streams has shown a strong positive effect of warming on primary production both among streams [@demars2011; @padfield2017] and within streams seasonally [@ogorman2012; @hood2018]. Invertebrates in these streams rely on autochthonous resources [@ogorman2012; @nelson2020], thus the dynamics of primary production have a strong control on consumer energy demand [@junker2020]. We, therefore, predicted that annual organic matter fluxes to consumers would increase with stream temperature mirroring patterns in resource availability and consumer energy demand. We also hypothesized that temperature would act as a principle environmental filter on community assembly and organic matter fluxes by favoring 'fast' life-history traits. Specifically, we predicted that warming temperatures among streams would lead to a decrease in average body size and an increase in average biomass turnover (i.e., *P:B* ratio) of populations, channeling the majority of energy flux through these taxa among stream communities. We also predicted that *within* communities, organic matter fluxes would be driven by small-bodied and high *P:B* taxa, and that these would not arise by random change [i.e., a 'selection effect' *sensu* @huston1997]. Our results should help refine general predictions about how warming, and its influence on key traits, is likely to shape energy flux through animal food webs.

@@

Annual community OM fluxes mirrored patterns of secondary production reported previously [@junker2020]. Organic matter flux to invertebrates varied ~`r round((max_flux$flux_mg_m_y_mean/min_flux$flux_mg_m_y_mean),0)`-fold among streams from `r round((min_flux$flux_mg_m_y_mean/1000), 1)`[`r round((min_flux$flux_mg_m_y_quant2.5/1000),1)` -- `r round((min_flux$flux_mg_m_y_quant97.5/1000),1)`] to `r round((max_flux$flux_mg_m_y_mean/1000),1)` [`r round((max_flux$flux_mg_m_y_quant2.5/1000),1)` -- `r round((max_flux$flux_mg_m_y_quant97.5/1000),1)`; mean [95% percentile interval (PI)]:] g AFDM $m^{-2} y^{-1}$) and was positively related to temperature (Figure 1A).

@@

In general, OM fluxes were unevenly distributed among taxa (Gini inequality coefficients ranged from `r gini_summary %>% select('Non.normalized.Gini_mean') %>% min %>% round(2)` [`r gini_summary %>% data.frame %>% dplyr::slice_min(Non.normalized.Gini_mean) %>% select(Non.normalized.Gini_quant2.5) %>% round(2)` -- `r gini_summary %>% data.frame %>% dplyr::slice_min(Non.normalized.Gini_mean) %>%  select('Non.normalized.Gini_quant97.5') %>% round(2)`, 95% PI] to `r gini_summary %>% data.frame %>% dplyr::slice_max(Non.normalized.Gini_mean) %>%  select('Non.normalized.Gini_mean') %>% round(2)` [`r gini_summary %>% data.frame %>% dplyr::slice_max(Non.normalized.Gini_mean) %>% select('Non.normalized.Gini_quant2.5') %>% round(2)` -- `r gini_summary %>% data.frame %>% dplyr::slice_max(Non.normalized.Gini_mean) %>% select('Non.normalized.Gini_quant97.5') %>% round(2)`, 95% PI; Table S1] and were dominated by insects in the families Simuliidae and Chironomidae, pulmonated snails (*Radix balthica*), and oligochaete worms (Figure 2A & B; Figure S2). In an absolute sense, ~85% of total OM flux was contributed by `r n_dom_min$n_dom` to `r n_dom_max$n_dom` taxa, which comprised only `r round(perc_dom_min$dom_spp*100,0)`% to `r round(perc_dom_max$dom_spp*100,0)`% of total taxon richness. Although differences in evenness among streams were partially attributed to variation in taxon richness (range: `r spp_richness %>% select(taxon_n) %>% min` to `r spp_richness %>% select(taxon_n) %>% max`), fluxes were still highly uneven after accounting for differences in richness (i.e., similar 'Normalized' Gini coefficients, Table S1). As average stream temperature increased, OM fluxes shifted from dominance by Simuliidae in the coolest stream to Chironomidae and *R. balthica* at moderate temperatures. In the warmest stream, where maximum temperatures approach ~40$^\circ$C, taxon richness was lowest and OM fluxes were dominated by oligochaete worms in the family Naididae, the chironomid *Cricotpous sylvestris*, and *R. balthica*. Among-stream differences in evenness and dominance of OM fluxes were not related to temperature (Appendix II Table S1).

@@

Within stream communities, there were diverse body size--OM flux relationships, with some streams showing skew toward larger-bodied taxa (positive skew), others with most OM fluxes through smaller-bodied sizes (negative skew), and some showed no difference among body sizes (skew range: `r m_skew_summ %>% slice_min(M_mg_ind_skew_quant50) %>% select(M_mg_ind_skew_quant50)%>% round(2)` to `r m_skew_summ %>% slice_max(M_mg_ind_skew_quant50) %>% select(M_mg_ind_skew_quant50) %>% round(2)`, Figure 3A & B). Skew estimates with body size (). Skew estimates of OM fluxes in relation to body size showed little association with stream temperature except in the warmest stream, where fluxes were heavily skewed toward small-bodied taxa. Similarly, skew in OM fluxes in relation to *P:B* ratios varied among streams, ranging from `r pb_skew_summ %>% slice_min(pb_y_skew_quant50) %>% select(pb_y_skew_quant50) %>% round(2)` to `r pb_skew_summ %>% slice_max(pb_y_skew_quant50) %>% select(pb_y_skew_quant50) %>% round(2)`. In this case, OM fluxes to consumers skewed increasingly toward higher turnover (*P:B*) taxa with increasing temperature (Figure 4A, B).

@@

While a growing body of theoretical and empirical study has enhanced our knowledge of temperature-mediated changes to ecosystems [e.g., @oconnor2009], general patterns are elusive and empirical studies often idiosyncratic [@zhang2017; @nelson2017], especially at higher levels of organization such as communities and food webs [@walther2002; @woodward2010]. We documented shifts in the pathways of material flows across a wide natural temperature gradient and found that increasing temperatures were associated with reductions in mean population body size and increases in mean population biomass turnover across stream invertebrate communities. In addition, we found that warming systematically skewed OM fluxes *within* communities such that, OM flows were increasingly dominated by rapid life-cycle taxa. Lastly, we discovered that the distribution of fluxes *within* communities was non-randomly organized, especially with respect to *P:B* ratios at moderate to high temperatures, suggesting that environmental filtering is especially important in the community assembly of warm communities. These patterns suggest that the acceleration of energy and material fluxes through ecosystems in both an absolute and relative sense may be a general effect of environmental warming. Further, environmental filtering that is driven by warming is likely governed by both the traits of species present and the ways in which energy and material fluxes are distributed among taxa.

Warming increased rates of materials flux in an absolute sense largely through shifts in community body-size structure. At the community level, we observed a ~`r ((exp(mean(pb_spp_temp_coefs, na.rm = TRUE)) - 1)*100) %>% round(0)`% increase in the mean biomass turnover rate (*P:B*; y^-1^) of populations for each 1$^\circ$C of warming. Although this effect among communities may be partially attributed to the thermodynamic influence of temperature [@gillooly2001], biomass turnover rate is also closely related to organism body size [@brown2004; @huryn2007]. Thus, warming may influence the distributions of *P:B* ratios directly through its effects on metabolic rate and indirectly through reductions in organism body size. The increase in *P:B* we observed across communities was similar in magnitude to the reduction in body size (i.e., +`r ((exp(mean(pb_spp_temp_coefs, na.rm = TRUE)) - 1)*100) %>% round(0)`% vs -`r ((exp(mean(m_spp_temp_coefs, na.rm = TRUE)) - 1)*100) %>% round(0)`%, respectively), reinforcing the fundamental connection between organism body size and *P:B*, but also suggesting increases in mean population *P:B* across communities were, may in large part, be attributed to reduced organism body sizes (Figure 1B & C). Organism body size is related to a number of other ecological attributes [@peters1983] and changes in body size with increasing temperatures are likely to have broad implications for ecosystems in a changing climate [@gibert2019a].

@@

A growing body of theoretical and empirical research is beginning to explore the implications of warming for the stability of ecosystems [@fussmann2014]. 'Natural laboratories', such as the geothermal watershed studied here, offer a unique opportunity to isolate the effects of warming and the dominant processes by which it may modulate ecosystem structure and dynamics [@ogorman2014]. Particularly relevant in our study is the stability and functioning of light-driven ecosystems in response to present and future warming. One aspect of climate change that is especially pronounced in high-latitude ecosystems is the decoupling of light regimes---and thereby energy supply regimes---and the timing of temperature-driven metabolic demands [@huryn2019a; @mcmeans2015]. Here, our within-community contrast in how warming skewed fluxes through *P:B* vs. body size taxa lends potential insight into the proximate ways in which temperature alters ecosystem structure and function. Temperature-size responses vary with a number of ecological factors [@ohlberger2013]; however, warming-induced body size reductions act as a stabilizing process in consumer-resource interactions by reducing consumer:resource biomass ratios [@osmond2017; @gilbert2014; @sentis2017] and consumer energy demand relative to resource supply [@kozlowski2004; @delong2012a; @mccann2011]. This gives credence to alternative models of organism body size optimization influenced by, for example, the balance of resource supply and energy demand [e.g., @kozlowski2004; @delong2012a]. Crucially, these alternative models highlight the trade-off between asymptotic body size and metabolic demand [@delong2012a], a trade-off that may be particularly important in seasonally variable energetic regimes. In the absence of other limiting factors, increasing temperatures are likely to make seasonal "boom and bust" cycles of primary production more extreme because of the interactive effects of light and temperature limitation on photosynthetic rates [e.g., @rae1998]. This process may increase the autocorrelation, as well as, the magnitude (i.e., enrichment) of resource supply, thereby, exaggerating the selection for reduced body sizes because larger and/or slower consumers may destabilize consumer-resource dynamics in the face of increasingly extreme seasonal autocorrelation of resource growth [@greyson-gaito2022]. This suggesting a connection between warming responses and a more general food web dynamic of increased relative rates of resource supply and consumer demand [e.g., paradox of enrichment, @mccann2011]. Therefore, the general response of reduced body size along climactic or geographic temperature gradients may reflect, in part, constraints imparted by food web dynamics, in addition to, physiological mechanisms [e.g., oxygen, @deutsch2022; differential responses of biological process rates, @forster2011]. The importance of this dynamic may be especially under-appreciated in high-latitude ecosystems [@mcmeans2020].

We documented the important role of temperature in structuring key functional triats and the relative distribution of material fluxes across natural stream temperature gradient. Higher temperatures were associated with increased total flux through the food web as well as reductions in average population body size and increases in population biomass turnover. Further, biomass turnover rate (but not body size) was an increasingly important trait for structuring OM fluxes at warmer temperatures. Our results lend support to the idea that warming may reduce organism size and also 'speed up' ecosystem dynamics in both an absolute and relative sense. These changes have important implications for the maintenance of biodiversity as well as for the connections between biodiversity and the magnitude and stability of ecosystem energy and material cycles in a warming world.

@@

### Construction zone

Possible other themes to layer in:

effects of temperature on species evenness, evenness of fluxes, and the statistical and ecological constraints on the organization of energy fluxes in communities (e.g., [@diaz2021a] SAD are more uneven than their statistical background and energy fluxes are also but there is important structure in this unevenness). This relates to the random vs non-random patterning questions...pull in the concepts of niche filtering and neutral processes in community assembly. Pull a little thinking from @saito2021.

Below is more of a set-up for the next paper but a little foreshadowing and hand waving may find a place in here:

These maximize energy intake when resources are available [@atkinson1997]

- Seasonal systems have unique constraints
- These seasonal constraints are greater at higher temps

This creates a situation where body sizes are selected for that 1) stabilize c-r dynamics [@osmond2017], ecological stability response to warming [@fussmann2014],...
2) optimize for energy availability and demand, 3) and this selection is stronger at extreme temperatures.

### Spare(d) words from previous versions. Keeping around for now.

directly kinetic effects and indirectly through the distribution of species traits, such as body size, within and across species [@bideault2019; @nelson2017].

environmental filtering of compared the observed distributions of material fluxes among consumer populations to random distributions to quantify the likelihood the patterning of organic matter fluxes arise through random ordering of species or  measure the potential strength of environmental filtering on the relative distribution of organic matter   we tested the potential strength of environmental filtering to determine the distribution of within stream energy fluxes by comparing the  how likely it was  Third, we hypothesized that the selective pressure from environmental temperature on body sizes and population turnover would be greatest at temperature extremes, therefore, the patterning of energy fluxes will be 'non-random' at coldest and warmest temperatures.

Yet, considerable uncertainty remains in the predicted short- and long-term effects of climate warming, especially at higher levels of organization such as communities and food webs [@woodward2010; @walther2002]. For example, species interactions in food webs may modulate species' TSRs [@delong2014], leading to varied responses among species and trophic levels [@ohlberger2013]. Additionally, warming-induced body size reductions may act as a stabilizing process in some consumer-resource interactions through reduced consumer:resource biomass ratios [@osmond2017] and reductions in consumer growth rates relative to resource supply [@kozlowski2004; @delong2012a; @mccann2011; @greyson-gaito2022].

For example, we lack a complete understanding of TSRs to explain deviations that exist across taxonomic groups (e.g., diatoms, @adams2013; invertebrates @zeuss2017; birds @geist1987; and fish @rypel2014). Explanations for these deviations vary [@ohlberger2013], some highlight additional processes such as competition and predation [@delong2014], while others propose alternative models of organism body size optimization, such as the balance of resource supply and energy demand [e.g., @kozlowski2004; @delong2012a]. Crucially, these alternative models highlight the trade-off between asymptotic body size and metabolic demand [@delong2012a] that can shape life history patterns (i.e., voltinism, @zeuss2017; @nelson2020b).  and community structure across ecosystems...

Recent, models have proposed  However, 
- both bottom-up (e.g. resource supply) and top-down (e.g., predation) proceesses
- competition and predation [@delong2014], stabilize c-r dynamics [@osmond2017], optimization, such as the balance of resource supply and energy demand [e.g., @kozlowski2004; @delong2012a]

>>>>

Submission journal: Ecology

Submission type: Article

Manuscript details:

- Abstract word count: 256/350

- Main text word count (w/ references): 5625

- Page count w/ references: 32/30

- Appendices: 2

Key words: climate change; energy flux; food webs; species traits; warming

Open Research Statement: All data and code used in this work are publicly available at the following link: https://github.com/jimjunker1/temperature_energy-flux.

@@

Warming temperatures are altering communities and trophic networks globally. While the influence of warming on food webs is often context-dependent, increasing temperatures are predicted to influence communities in two fundamental ways: 1) by reducing average body size and 2) by increasing individual metabolic rates. These warming-induced changes have the potential to influence the distribution of food web fluxes, food web stability, and the importance of environmental and biological factors in community assembly. Here, we quantified patterns and the relative distribution of organic matter fluxes through stream food webs spanning a natural ~25$^\circ$C temperature gradient. We then related these patterns to species- and community-level trait distributions (body size, *M* and biomass turnover, *P:B*) within and across ecosystems. We predicted that 1) communities in warmer streams would skew towards smaller *M* and higher *P:B* and 2) organic matter fluxes within warmer communities would increasingly skew towards smaller, higher *P:B* populations. Across the temperature gradient, communities in warmer streams were indeed composed of smaller-bodied and higher biomass-turnover populations on average. Additionally, material fluxes within warmer streams were increasingly skewed towards high-turnover populations, demonstrating that warming can restructure organic matter fluxes in both an absolute and relative sense. With warming, the relative distribution of organic matter fluxes also appeared to be increasingly 'non-random', suggesting the potential for stronger selection for these traits with increasing temperature. Our study suggests that a warming world will be 'smaller and faster', and that warming has the potential to shape pathways of energy and material flows in food webs in ways that may be predictable.

@@

<!--- warming and effects in food webs --->
<!--- Speeding up of food web: Species levels--->
Increasing global temperatures can influence the provision and maintenance of ecosystem services by modifying the network of species interactions that underpin ecosystem functions [@brose2012; @woodward2010; @deruiter1995; @thompson2012]. The effects of warming permeate across levels of biological organization from its control on individual metabolic rates [@gillooly2001; @brown2004] and biological activity (e.g., attack rate, handling time, growth rates, etc.; @dell2014, to broad-scale shifts in community assembly and structure [@saito2021; @gibert2019a; @nelson2017a]. Across global climate gradients, these temperature-induced changes have the potential to mediate food web stability [@baiser2019] by altering the acquisition and allocation of resources [@zhang2017] and the magnitudes and relative distribution of energy fluxes in food webs [@may1972; @mccann1998; @barnes2018].

Understanding how warming alters energy fluxes in food webs requires information about how temperature modifies connections between ecosystem structure (i.e. biodiversity) and function. Such connections hinge upon how warming influences functional trait distributions and how these functional traits translate to energy demands [@loreau2001; @norberg2001; @norberg2004]. Additionally, warming may modify both deterministic [e.g., environmental/niche filtering, @whittaker1962] and stochastic [e.g., neutral theory, @hubbell2001] community assembly processes [@saito2021], leading to many possible relationships between warming and species' abundance and trait distributions. For example, although temperature appears to have minimal control on species richness across taxonomic groups [e.g., @bastazini2021], warming may have profound effects on the dominance structure (i.e. evenness) of ecological communities, often by reducing community evenness and favoring a few well-adapted species [@hillebrand2008]. This strong environmental filtering is likely to skew trait distributions, to some extent, in natural communities [@therriault1999]. However, the relative distribution of traits can be modified by additional processes beyond environmental filtering that alter species relative abundance distributions [e.g., species interactions, @therriault1999; demographic stochasticity, @hubbell2001]. These additional processes can exaggerate or counter any skew in trait distribution imparted through environmental filtering and may modify the relationship between species' traits and the absolute and relative energy demands in food webs.

<!--- Connect traits to the 'speed' of ecosystem energy fluxes--->
Body size is a fundamental trait that is strongly influenced by temperature and has great potential to influence energy flux [@atkinson1994; @gardner2011; @daufresne2009]. Body size reductions arise from multiple mechanisms acting at different levels of organization, including the increased relative abundance of smaller species within warmer communities [@bergmann1848] or smaller individuals in warmer populations [@james1970], or reduced absolute body size of warmer individuals [@atkinson1994]. These temperature-size relationships are likely to influence the distribution of body sizes both across and within ecosystems, such that warmer communities should contain smaller species and populations, on average, and smaller organisms should be more dominant within a community. These changes can have important implications for ecosystems pattern and processes that control energy and material fluxes in ecosystems because body size is a strong determinant of species life history patterns [@peters1983; @zeuss2017; @altermatt2010; @nelson2020b] and, developmental [@angilletta2004] or metabolic rates [@brown2004; @gillooly2001].

<!--- Write the main goals of the paragraph--->
Metabolic rate, in addition to being controlled by body size, is another key trait that is influence by temperature and tied to variation in energy flux through food webs [@gillooly2001]. Warming can influence metabolic rates indirectly through reductions in body size, as well as directly through its effects on subcellular kinetics [@bideault2019; @osmond2017]. These processes have been shown to modify ecosystem patterns through changes in population abundance [@bernhardt2018], consumer-resource interactions [@bideault2019], and food web structure [@gibert2019a]. Importantly, the kinetic effects of temperature are tied to many biological processes [@dell2014], among them growth rate [@gillooly2001], developmental rate [@zuo2012; @nelson2020b], voltinism [@zeuss2017], and population turnover rate [@huryn2007; @brown2004]. Taken together, these changes suggest that warming should lead to a smaller, faster world in which ecosystem processes 'speed up' through the effects of smaller body size and higher turnover rates, with a potentially strong imprint on food web structure and energy fluxes [@gibert2019a].

We quantified patterns of organic matter flux in stream food webs across a natural temperature gradient (\~5--28$^\circ$C) in southwestern Iceland. Previous research in these streams has shown a strong positive effect of warming on primary production both among streams [@demars2011; @padfield2017] and within streams seasonally [@ogorman2012; @hood2018]. Invertebrates in these streams rely on autochthonous resources [@ogorman2012; @nelson2020], thus the dynamics of primary production have a strong control on consumer energy demand [@junker2020]. We, therefore, predicted that annual organic matter fluxes to consumers would increase with stream temperature mirroring patterns in resource availability and consumer energy demand. We also hypothesized that temperature would act as a principle environmental filter on community assembly and organic matter fluxes by favoring 'fast' life-history traits. Specifically, we predicted that warming temperatures among streams would lead to a decrease in average body size and an increase in average biomass turnover (i.e., *P:B* ratio) of populations. We also predicted that *within* communities, organic matter fluxes would be dominated by small-bodied and high *P:B* taxa at higher temperatures and that these patterns would not arise by random change [i.e., a 'selection effect' *sensu* @huston1997]. Our results should help refine general predictions about how warming, and its influence on key traits, is likely to shape energy flux through animal food webs.

@@

Annual community OM fluxes mirrored patterns of secondary production reported previously [@junker2020]. Organic matter flux to invertebrates varied ~`r round((max_flux$flux_mg_m_y_mean/min_flux$flux_mg_m_y_mean),0)`-fold among streams from `r round((min_flux$flux_mg_m_y_mean/1000), 1)` (`r round((min_flux$flux_mg_m_y_quant2.5/1000),1)` -- `r round((min_flux$flux_mg_m_y_quant97.5/1000),1)`; mean [95% percentile interval (PI)]) to `r round((max_flux$flux_mg_m_y_mean/1000),1)` (`r round((max_flux$flux_mg_m_y_quant2.5/1000),1)` -- `r round((max_flux$flux_mg_m_y_quant97.5/1000),1)`) g AFDM $m^{-2} y^{-1}$) and was positively related to temperature (Figure 1A).

@@

In general, OM fluxes were unevenly distributed among taxa (Gini inequality coefficients ranged from `r gini_summary %>% select('Non.normalized.Gini_mean') %>% min %>% round(2)` (`r gini_summary %>% data.frame %>% dplyr::slice_min(Non.normalized.Gini_mean) %>% select(Non.normalized.Gini_quant2.5) %>% round(2)` -- `r gini_summary %>% data.frame %>% dplyr::slice_min(Non.normalized.Gini_mean) %>%  select('Non.normalized.Gini_quant97.5') %>% round(2)`, 95% PI) to `r gini_summary %>% data.frame %>% dplyr::slice_max(Non.normalized.Gini_mean) %>%  select('Non.normalized.Gini_mean') %>% round(2)` (`r gini_summary %>% data.frame %>% dplyr::slice_max(Non.normalized.Gini_mean) %>% select('Non.normalized.Gini_quant2.5') %>% round(2)` -- `r gini_summary %>% data.frame %>% dplyr::slice_max(Non.normalized.Gini_mean) %>% select('Non.normalized.Gini_quant97.5') %>% round(2)`, 95% PI; Table S1) and were dominated by insects in the families Simuliidae and Chironomidae, pulmonated snails (*Radix balthica*), and oligochaete worms (Figure 2A & B; Figure S2). In an absolute sense, ~85% of total OM flux was contributed by `r n_dom_min$n_dom` to `r n_dom_max$n_dom` taxa, which comprised only `r round(perc_dom_min$dom_spp*100,0)`% to `r round(perc_dom_max$dom_spp*100,0)`% of total taxon richness. Although differences in evenness among streams were partially attributed to variation in taxon richness (range: `r spp_richness %>% select(taxon_n) %>% min` to `r spp_richness %>% select(taxon_n) %>% max`), fluxes were still highly uneven after accounting for differences in richness (i.e., similar 'Normalized' Gini coefficients, Table S1). As average stream temperature increased, OM fluxes shifted from dominance by Simuliidae in the coolest stream to Chironomidae and *R. balthica* at moderate temperatures. In the warmest stream, where maximum temperatures approach ~40$^\circ$C, taxon richness was lowest and OM fluxes were dominated by oligochaete worms in the family Naididae, the chironomid *Cricotpous sylvestris*, and *R. balthica*. Among-stream differences in evenness and dominance of OM fluxes were not related to temperature (Appendix II Table S1).

@@

Within stream communities, there were diverse body size--OM flux relationships, with some streams showing skew toward larger-bodied taxa (positive skew), others with most OM fluxes through smaller-bodied sizes (negative skew), and some showed no difference among body sizes (skew range: `r m_skew_summ %>% slice_min(M_mg_ind_skew_quant50) %>% select(M_mg_ind_skew_quant50)%>% round(2)` to `r m_skew_summ %>% slice_max(M_mg_ind_skew_quant50) %>% select(M_mg_ind_skew_quant50) %>% round(2)`, Figure 3A & B). Skew estimates of OM fluxes in relation to body size showed little association with stream temperature except in the warmest stream, where fluxes were heavily skewed toward small-bodied taxa. Similarly, skew in OM fluxes in relation to *P:B* ratios varied among streams, ranging from `r pb_skew_summ %>% slice_min(pb_y_skew_quant50) %>% select(pb_y_skew_quant50) %>% round(2)` to `r pb_skew_summ %>% slice_max(pb_y_skew_quant50) %>% select(pb_y_skew_quant50) %>% round(2)`. In this case, OM fluxes to consumers skewed increasingly toward higher turnover (*P:B*) taxa with increasing temperature (Figure 4A, B).

@@

While a growing body of theoretical and empirical study has enhanced our knowledge of temperature-mediated changes to ecosystems [e.g., @oconnor2009], general patterns are elusive and empirical studies often idiosyncratic [@zhang2017; @nelson2017], especially at higher levels of organization such as communities and food webs [@walther2002; @woodward2010]. We documented shifts in the pathways of material flows across a wide natural temperature gradient and found that increasing temperatures were associated with reductions in mean population body size and increases in mean population biomass turnover across stream invertebrate communities. In addition, we found that warming systematically skewed OM fluxes *within* communities such that, OM flows were increasingly dominated by rapid life-cycle taxa. Lastly, we discovered that the distribution of fluxes within communities was non-randomly organized, especially with respect to *P:B* ratios at moderate to high temperatures, suggesting that temperature was especially important in structuring the relative performance and resource acquisition among taxa within warm communities. These patterns suggest that the acceleration of energy and material fluxes through ecosystems in both an absolute and relative sense may be a general effect of environmental warming. Further, environmental filtering that is driven by warming is likely governed by both the traits of species present and the ways in which energy and material fluxes are distributed among taxa.

Warming increased rates of materials flux in an absolute sense largely through shifts in community body-size structure. At the community level, we observed a ~`r ((exp(mean(pb_spp_temp_coefs, na.rm = TRUE)) - 1)*100) %>% round(0)`% increase in the mean biomass turnover rate (*P:B*; y^-1^) of populations for each 1$^\circ$C of warming. Although this effect among communities may be partially attributed to the thermodynamic influence of temperature [@gillooly2001], biomass turnover rate is also closely related to organism body size [@brown2004; @huryn2007]. Thus, warming may influence the distributions of *P:B* ratios directly through its effects on metabolic rate and indirectly through reductions in organism body size. The increase in *P:B* we observed across communities was similar in magnitude to the reduction in body size (i.e., +`r ((exp(mean(pb_spp_temp_coefs, na.rm = TRUE)) - 1)*100) %>% round(0)`% vs -`r ((exp(mean(m_spp_temp_coefs, na.rm = TRUE)) - 1)*100) %>% round(0)`%, respectively), reinforcing the fundamental connection between organism body size and *P:B*, but also suggesting increases in mean population *P:B* across communities may, in large part, be attributed to reduced organism body sizes (Figure 1B & C). Organism body size is related to a number of other ecological attributes [@peters1983] and changes in body size with increasing temperatures are likely to have broad implications for ecosystems in a changing climate [@gibert2019a].

@@

A growing body of theoretical and empirical research is beginning to explore the implications of warming for the stability of ecosystems [@fussmann2014]. 'Natural laboratories', such as the geothermal watershed studied here, offer a unique opportunity to isolate the effects of warming and the dominant processes by which it may modulate ecosystem structure and dynamics [@ogorman2014]. Particularly relevant in our study is the stability and functioning of light-driven ecosystems in response to present and future warming. One aspect of climate change that is especially pronounced in high-latitude ecosystems is the decoupling of light regimes---and thereby energy supply regimes---and the timing of temperature-driven metabolic demands [@huryn2019a; @mcmeans2015]. Here, our within-community contrast in how warming skewed fluxes through taxa with respect to *P:B* vs. body size may shed light on the proximate ways in which temperature alters ecosystem structure and function. Temperature--size responses vary with a number of ecological factors [@ohlberger2013]; however, warming-induced body size reductions act as a stabilizing process in consumer-resource interactions by reducing consumer:resource biomass ratios [@osmond2017; @gilbert2014; @sentis2017] and consumer energy demand relative to resource supply [@kozlowski2004; @delong2012a; @mccann2011]. This gives credence to alternative models of organism body size optimization influenced by, for example, the balance of resource supply and energy demand [e.g., @kozlowski2004; @delong2012a]. Crucially, these alternative models highlight the trade-off between asymptotic body size and metabolic demand [@delong2012a], a trade-off that may be particularly important in seasonally variable energetic regimes. In the absence of other limiting factors, increasing temperatures are likely to make seasonal "boom and bust" cycles of primary production more extreme because of the interactive effects of light and temperature limitation on photosynthetic rates [e.g., @rae1998]. This process may increase the autocorrelation, as well as, the magnitude (i.e., enrichment) of resource supply, thereby, exaggerating the selection for reduced body sizes because larger and/or slower consumers may destabilize consumer-resource dynamics in the face of increasingly extreme seasonal autocorrelation of resource growth [@greyson-gaito2022]. Thus, suggesting a connection between warming responses and a more general food web dynamic of increased relative rates of resource supply and consumer demand [e.g., paradox of enrichment, @mccann2011]. Therefore, the general response of reduced body size along climactic or geographic temperature gradients may reflect, in part, constraints imparted by food web dynamics, in addition to, physiological mechanisms [e.g., oxygen, @deutsch2022; differential responses of biological process rates, @forster2011]. The importance of this dynamic may be especially under-appreciated in high-latitude ecosystems [@mcmeans2020].

We documented the important role of temperature in structuring key functional traits and the relative distribution of material fluxes across natural stream temperature gradient. Higher temperatures were associated with increased total flux through the food web as well as reductions in average population body size and increases in population biomass turnover. Further, biomass turnover rate (but not body size) was an increasingly important trait for structuring OM fluxes at warmer temperatures. Our results lend support to the idea that warming may reduce organism size and also 'speed up' ecosystem dynamics in both an absolute and relative sense. These changes have important implications for the maintenance of biodiversity as well as for the connections between biodiversity and the magnitude and stability of ecosystem energy and material cycles in a warming world.

@@
